---
title: "Extract latent factor scores for continuous measures of anxiety"
author: "Genevieve Morneau-Vaillancourt"
output: html_document
---

Load packages (install packages first by removing '#' if needed)
```{r}
# install.packages('lavaan')
library(lavaan)

# install.packages('tidyverse')
library(tidyverse)

#install.packages('psych')
library(psych)
```

Read your data
Different extensions can be read, for example .txt; .sav; .xlsx
Some extensions including .sav will necessitate packages such as haven or foreign
For more information, see tutorials such as: https://www.datafiles.samhsa.gov/get-help/format-specific-issues/how-do-i-read-data-r
```{r}
# line of code to upload your data e.g. datafile <- read_delim("filename.txt")
```

Run lavaan model including ≥3 time points using maximum likelihood for missing data, assuming no correlations between residuals.
Note: Replace var1, var2, var3 with the names of your variables and add additional variables if you have more than 3 time points. Replace datafile with the label for your uploaded datafile
The default name of the factor score variable in the model below is "latent_factor" 
```{r}
# Create model to estimate a latent factor 
model <- '
latent_factor =~ var1 + var2 + var3
'

# Fit model
fitmodel <- sem(model, missing = "FIML", estimator = "ML", data = datafile)
summary(fitmodel)

# Save coefficients
factor_scores <- predict(fitmodel)
```

Extract latent factor scores
```{r}
# Create and fit a new model including ID variables - this step is to be able to merge extracted latent factor scores with ID variables. Please ignore warning messages after fitting the model as model fit is irrelevant here; it's a hack to get the IDs in the lavaan internal data matrix.
model_id <- '
participant_id ~~ var1 + var2 + var3
'
fitmodel_id <- sem(model_id, missing = "FIML", estimator = "ML", data = datafile)

# Create a new dataset including the saved factor scores. Confirm that the id variable is the first column in the internal data matrix e.g. colnames(fitmodel_id@Data@X[[1]])
datafile2 <- fitmodel_id@Data@X[[1]]

# Merge datasets including estimated factor scores and ID variables
datafile3 <- data.frame(factor_scores, participant_id = datafile2[,1])
```

Merge with original datafile
```{r}
datafile_final <- merge(datafile, datafile3, by = c("participant_id"), all.x = TRUE)
```

Descriptives of latent factor score using e.g. psych package or base r functions
```{r latent score descriptives}
psych::describe(datafile_final$latent_factor)
```

Standardise latent factor score
```{r stdise latent factor}
datafile_final <- datafile_final %>%
  mutate(latent_fs_std = as.numeric(scale(latent_factor)))

psych::describe(datafile_final$latent_fs_std)
```

You can then save this new datafile and perform the GWAS on "latent_fs_std".
```{r}
# code to save your datafile
datafile_final <- datafile_final %>%
  mutate(FID = participant_id, IID = participant_id)

write_tsv(datafile_final %>% select(FID, IID, latent_fs_std), 
          "quantanx_pheno.txt")
```
