---
title: "Quantitative anxiety symptom genome-wide meta-analysis"
author: "Megan Skelton"
output: html_document
---

Analysis was run on the KCL CREATE HPC: King's College London. (2022). King's Computational Research, Engineering and Technology Environment (CREATE). Retrieved March 2, 2022, from https://doi.org/10.18742/rnvf-m076

# Meta-analysis
```{bash METAL script}

echo '
COLUMNCOUNTING   STRICT
SCHEME           STDERR
AVERAGEFREQ      ON
MINMAXFREQ       ON
CUSTOMVARIABLE   N
GENOMICCONTROL   OFF
OVERLAP	    OFF
ADDFILTER EAF > 0.01
ADDFILTER INFO > 0.6

MARKERLABEL      MARKER
ALLELELABELS     EA NEA           
EFFECTLABEL      BETA
PVALUELABEL      P
STDERR           SE
FREQLABEL        EAF
WEIGHTLABEL      N

PROCESSFILE /qcd_cohort_sumstats/AGDS_all.QCed
PROCESSFILE /qcd_cohort_sumstats/ALSPAC_GAD7_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/EstBB_ESTQ2_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/GenScot_GHQ_ANX_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/GLAD_UKB_NBR_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/TEDS_GAD10_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/PsyCoLaus_STAI_STATE_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/MOBA_GAD7_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/MEGA_STAI_TRAIT_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/MVP_GAD2_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/PROTECTAD_140624_GAD7_noX.QCed
PROCESSFILE /qcd_cohort_sumstats/lifelines_GSA_GAD_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/lifelines_CYTO_GAD_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/lifelines_AFFY_GAD_1209_Final.QCed
PROCESSFILE /qcd_cohort_sumstats/TRAILS_allchr.QCed

OUTFILE quantanx_metaanalysis .tbl
ANALYZE HETEROGENEITY
VERBOSE ON 
QUIT
' > gadsympt_METAL.txt
module load metal/2020-05-05-gcc-13.2.0
metal gadsympt_METAL.txt

```

## Post-MA QC
Add SNP (rsID) to output by merging with reference panel e.g. HRC, joining on Marker (CHR:POS), and add CHR BP or split Marker for separate columns.
Restrict to SNPs >50% cohorts contributed to and with EAF 0.01 < Freq1 < 0.99. 
Select required columns and rename. Order by CHR:POS.
```{bash}

# check where CHR, BP, SNP are in your REF file, and where MARKERID is that you're merging on - here it was $1 
awk -F'\t' -v OFS='\t' 'NR==FNR {
    CHR[$1] = $18
    BP[$1]  = $19
    SNP[$1] = $17
    next
}
FNR==1 {
    print $0, "CHR", "BP", "SNP"
    next
}
{
    print $0, CHR[$1], BP[$1], SNP[$1]
}' \
    <(gunzip -c ref_file_with_CHR_BP_SNP) \
    <(gunzip -c quantanx_metaanalysis.gz) | gzip \
    > quantanx_metaanalysis_rsid.gz


gunzip -c quantanx_metaanalysis_rsid.gz \
| awk '
NR==1 { print; next }

{
    chr = $17 #assuming chr is column 17
    n_contrib = $14+1 # assuming HetDf (N cohorts - 1) is column 14, add 1 for N contributing cohorts

    if ((chr == 23 && n_contrib >= 4) || (chr != 23 && n_contrib >= 8)) { #different thresholds as subset provided x-chr data
        print
    }
}
' | gzip > quantanx_metaanalysis_filtered.gz

# Select the required cols
zcat quantanx_metaanalysis_filtered.gz | awk 'NR==1 {print "SNP", "MARKERID", "CHR", "BP", "A1", "A2", "FREQ", "BETA", "SE", "P", "N"; next}
{print $19, $1, $17, $18, $2, $3, $4, $8, $9, $10, $16}
' | sort -k3,3n -k4,4n | gzip > pgc-anx2-gadsymptsquant-2026_eur.txt.gz

```

# LDSC SNP h2
```{bash}

/ldsc/munge_sumstats.py \
--chunksize 50000 \
--snp SNP \
--sumstats /metaanalysis_sumstats/quantanx_metaanalysis_filtered.gz \
--out /munged/quantanx_metaanalysis \
--merge-alleles /ldsc/w_hm3.snplist

/ldsc/ldsc.py \
--h2 /munged/quantanx_metaanalysis.sumstats.gz \
--ref-ld-chr /ldsc/eur_w_ld_chr/ \
--w-ld-chr /ldsc/eur_w_ld_chr/ \
--out /snph2/quantanx_metaanalysis_ldsc_h2

```

# Manhattan plot and QQ plot
```{r}

library(data.table)
library(tidyverse)
library(qqman)

palette = c("#3d7667", "#abd4c9")

results <- fread("pgc-anx2-gadsymptsquant-2026_eur.txt.gz", header=T)

bitmap("quantanx_manhattan.png",
type = "png16m", height = 10, width = 30, res = 600)

par(cex.axis = 0.9, cex.lab = 1.3)
manhattan(results, 
chr="CHR", 
bp="BP",
p="P",
snp="SNP",
ylim = c(0, 16),
col = palette,
cex = 0.5,
suggestiveline = F,
genomewideline = -log10(5e-08))

dev.off()

bitmap("quantanx_qqplot.png", type = "png16m", 
height = 10, width = 10, res = 600)

par(mar=c(5,5,1,1))
qq(results$P,
cex.axis = 1.1,
cex.lab = 1.3,
xlim = c(0, 9),
ylim = c(0, 15))

dev.off()

```

# Significant loci
Using the FUMA output (as performed 2 rounds of clumping) from the following parameters:
r2 threshold 0.1 
second r2 threshold 0.05
LD blocks on locus 500kb
1000G Phase3 EUR
```{r}

fuma_snps <- fread("../FUMA_job686083/IndSigSNPs.txt")

loci_table <- fuma_snps %>%
  select(Locus = GenomicLocus,
         SNP = rsID,
         CHR = chr,
         BP = pos,
         p,
         NSigSNPs = nGWASSNPs)

sumstats <- fread("pgc-anx2-gadsymptsquant-2026_eur.txt.gz")

sumstats <- sumstats %>%
  select(SNP, A1, A2, A1Freq = FREQ, Beta = BETA, SE, N)

loci_table <- loci_table %>%
  left_join(., sumstats, by = "SNP") %>% 
  select(Locus, SNP, CHR, BP, A1, A2, A1Freq, Beta, SE, p, N, NSigSNPs)

all_snps <- fread("../FUMA_job686083/snps.txt")

all_snps_info <- all_snps %>%
  select(SNP = rsID, CHR = chr, nearestGene, dist) %>%
  left_join(loci_table, ., by = c("SNP", "CHR"))

```

# Novelty and replication
## LDTrait 
LDlink results using an r2 of 0.1 and a 500kb basepair window ref panel, saved in a txt file
```{r load ldtrait results}

ldtrait_quantanx <- fread("../ma_results/correction_aug25/ldlink_quantanx_r201_500kb_020925.txt")

```

### Internalising
Identify SNPs identified in previous GWAS of internalising traits
```{r identify replicated snps}

snps_internalising <- ldtrait_quantanx %>% filter(
  grepl("anx", `GWAS Trait`, ignore.case = T) | 
  grepl("depr", `GWAS Trait`, ignore.case = T) &
  !grepl("Bipolar disorder vs major depressive disorder \\(ordinary least squares \\(OLS\\)\\)", 
           `GWAS Trait`, ignore.case = TRUE) | #ignore this label captured by depr but is a comparison of bipolar + depr
  grepl("neuroti", `GWAS Trait`, ignore.case = T) |
  grepl("mdd", `GWAS Trait`, ignore.case = T) |
  grepl("worry", `GWAS Trait`, ignore.case = T) | 
  grepl("well-being", `GWAS Trait`, ignore.case = T) | 
  grepl("satisfaction", `GWAS Trait`, ignore.case = T) |
  grepl("feeling", `GWAS Trait`, ignore.case = T) |  #captures Feeling miserable, Feeling tense, Feeling nervous, Feeling worry, Feeling fed-up, Feeling guilty - confirmed the few that appear for this latter also have other internalising associations
  grepl("Post-traumatic stress", `GWAS Trait`, ignore.case = T) |
  `GWAS Trait` == "Positive affect" |
  `GWAS Trait` == "Suffering from nerves" | 
  `GWAS Trait` == "Major mood disorders" | # only one for this and also appeared for other internalising traits e.g. worry, neuroticism, depression (if had only been mood would want to check this wasn't driven by bip/mania)
  `GWAS Trait` == "Irritable mood" | # confirmed that all rsIDs for this also appeared for other internalising traits
  `GWAS Trait` == "Neurociticism") # capture this trait label which has a typo

nrow(snps_internalising)

# list the snps from our sumstats (query) that had internalising hits
snps_internalising_list <- snps_internalising %>%
  distinct(Query) %>%
  pull()

# create a binary variable to flag if locus had an internalising hit
loci_table <- loci_table %>%
  mutate(
    internalising_hit = if_else(SNP %in% snps_internalising_list, 1, 0)
  )

```

### Anxiety
```{r}

snps_anx_list <- snps_internalising %>% 
  filter(
    grepl("anx", `GWAS Trait`, ignore.case = T)) %>% 
  distinct(Query) %>%
  pull()

loci_table <- loci_table %>%
  mutate(
    anx_hit = if_else(SNP %in% snps_anx_list, 1, 0)
  )

loci_table_replications <- loci_table %>%
  select(Locus, SNP, CHR, BP, internalising_hit, anx_hit)

novel_snps <- setdiff(fuma_snps$rsID, snps_internalising$Query) # Novel = novel for internalising
length(novel_snps)

```

## Bedtools
Use FUMA to clump (same parameters as our sumstats) and bedtools to identify overlap with relevant sumstats not in GWAS Catalog

Li et al. anxiety
Friligkou et al. anxiety
Strom et al. anxiety
Adams et al. MDD

```{bash create BED files for bedtools}

for i in quantanx_GenomicRiskLoci.txt Stromanxiety_GenomicRiskLoci.txt Lianxiety_GenomicRiskLoci.txt Friligkouanxiety_GenomicRiskLoci.txt AdamsMDD23andme_GenomicRiskLoci.txt 
do
awk 'NR>1 {print "chr"$4, $7-1, $8, $1}' OFS='\t' "$i" > "${i}.BED"
done

module load bedtools2/2.31.0-gcc-12.3.0-python-3.11.6

for i in Stromanxiety_GenomicRiskLoci.txt.BED Lianxiety_GenomicRiskLoci.txt.BED Friligkouanxiety_GenomicRiskLoci.txt.BED AdamsMDD23andme_GenomicRiskLoci.txt.BED
do
bedtools intersect -a quantanx_GenomicRiskLoci.txt.BED -b "$i" > "intersect_quantanx_${i}.BED"
done

```

Merge individual results in R file
```{r}

library(data.table)
library(tidyverse)

quantanx_loci <- fread("quantanx_GenomicRiskLoci.txt", data.table = F)
strom <- fread("intersect_quantanx_Stromanxiety_GenomicRiskLoci.txt.BED.BED", data.table = F)
li <- fread("intersect_quantanx_Lianxiety_GenomicRiskLoci.txt.BED.BED", data.table = F)
adams <- fread("intersect_quantanx_AdamsMDD23andme_GenomicRiskLoci.txt.BED.BED", data.table = F)
friligkou <- fread("intersect_quantanx_Friligkouanxiety_GenomicRiskLoci.txt.BED.BED", data.table = F)

quantanx_loci <- quantanx_loci %>%
  select(GenomicLocus, chr, pos, start, end, rsID, IndSigSNPs, LeadSNPs) %>%
  rename(locus = GenomicLocus)

sumstat_names <- c("strom", "li", "friligkou", "adams")

for (name in sumstat_names) {
  df <- get(name)

  df <- df %>%
    rename(chr = V1, start = V2, end = V3, locus = V4) %>%
    select(locus) %>%
    mutate(!!paste0(name, "_locus") := locus)

  quantanx_loci <- left_join(quantanx_loci, df, by = "locus")
}

strom_loci <- fread("Stromanxiety_GenomicRiskLoci.txt", data.table = F)
li_loci <- fread("Lianxiety_GenomicRiskLoci.txt", data.table = F)
friligkou_loci <- fread("Friligkouanxiety_GenomicRiskLoci.txt", data.table = F)
adams_loci <- fread("AdamsMDD23andme_GenomicRiskLoci.txt", data.table = F)

sumstats_list <- list(
  strom_rsid = strom_loci,
  li_rsid = li_loci,
  friligkou_rsid = friligkou_loci,
  adams_rsid = adams_loci
)

quantanx_loci <- reduce(names(sumstats_list), function(quantanx_loci, sumstats_name) {
  sumstats_df <- sumstats_list[[sumstats_name]]
  
  quantanx_loci %>%
    mutate(!!sumstats_name := pmap_chr(list(chr, start, end), ~ {
      match_row <- sumstats_df %>%
        filter(
          chr == ..1 & (
            (pos >= ..2 & pos <= ..3) |   # SNP within locus
            (start <= ..3 & end >= ..2)   # reverse-matching where sumstats locus contains the quantanx region including any overlap, not just full containment
          )
        )
      if (nrow(match_row) > 0) match_row$rsID[1] else NA_character_ # takes first match, check for multi matches though for replication flag not critical, or replace match_row$rsID[1] with paste(match_row$rsID, collapse = ";")
    }))
}, 
.init = quantanx_loci)

# Split into 4 pairwise files of our trait rsID and comparison rsID to check R2 in PLINK
rep_cols <- c("strom_rsid", "li_rsid", "friligkou_rsid", "adams_rsid")

snp_pairs_by_study <- quantanx_loci %>%
  select(locus, rsID, all_of(rep_cols)) %>%
  pivot_longer(cols = all_of(rep_cols), names_to = "study", values_to = "replication_rsid") %>%
  separate_rows(replication_rsid, sep = ";", convert = TRUE) %>%
  transmute(
    study = str_remove(study, "_rsid"),
    SNP_A = rsID,
    SNP_B = replication_rsid
  )

snp_pairs_by_study %>%
  group_by(study) %>%
  group_split() %>%
  walk(~ write_tsv(.x %>% select(SNP_A, SNP_B),
                   file = paste0("compare_snp_pairs_", unique(.x$study), ".txt"),
                   col_names = FALSE,
                   na = ""))

```

```{bash}

for f in compare_snp_pairs_*.txt; do
  study=$(basename "$f" .txt | cut -d'_' -f4)
  echo "Running LD for $study..."
  while read -r a b; do
    # skip lines where b is empty (no SNP for comparison study in same locus)
    [ -z "$b" ] && continue
    plink \
      --bfile /users/k1756554/1kg_refpanels/g1000_eur \
      --ld $a $b \
      --out results_ld/${study}_${a}_${b}
  done < "$f"
done

```


```{r novelty flag}

ld_files <- list.files(path = "results_ld", pattern = "\\.log$", full.names = TRUE)

# define function to extract required information from log file
read_ld_log <- function(file){
  snps <- str_extract_all(basename(file), "rs[0-9]+")[[1]]
  lines <- readLines(file)
  r2_line <- lines[grepl("R-sq =", lines)]
  if(length(r2_line) == 0) return(NULL)
  r2 <- as.numeric(str_match(r2_line, "R-sq = ([0-9\\.eE+-]+)")[,2])
  tibble(
    SNP_A = snps[1],
    SNP_B = snps[2],
    R2 = r2
  )
}

ld_table <- map_dfr(ld_files, read_ld_log)
ld_table <- ld_table %>%
  distinct(SNP_A, SNP_B, .keep_all = TRUE) 

replication <- snp_pairs_by_study %>%
  left_join(ld_table, by = c("SNP_A", "SNP_B")) %>%
  mutate(replication_type = case_when(
    SNP_A == SNP_B ~ "Exact replication",
    !is.na(R2) & R2 >= 0.1 ~ "LD replication",
    TRUE ~ "Novel" #i.e. R2 <0.1 is considered novel
  ))
         
```

# MAGMA 
## Gene-based association test

```{r}

magma_genes <- fread("/FUMA_job686083/magma.genes.out.txt")
magma_genes_sig <- magma_genes %>% 
  filter(P < 2.506e-6) 
nrow(magma_genes_sig)
magma_genes_sig <- magma_genes_sig %>%
  mutate(CHR = as.integer(CHR)) %>%
  arrange(CHR, START)
write.table(magma_genes_sig, "magma_genes_sig.txt", 
              sep = "\t", 
              row.names = FALSE, 
              col.names = TRUE,
              quote = FALSE)

```

Clump all GWAS variants to p = 1 (i.e. every variant gets put in a clump) with clump-range flag and a gene regions file in the format specified here - https://www.cog-genomics.org/plink/1.9/resources#genelist
Contains one gene per row, with the following four columns:
Chromosome code (example file shows 1, 2 etc. not chr1, chr2)
Start of gene (base-pair units, 1-based)
End of gene (this position is included in the interval)
Gene ID
```{bash}

awk 'NR>1 {print $2, $3, $4, $10}' OFS='\t' magma_genes_sig.txt > magma_genes_sig_plinkformat.txt

plink --bfile /1kg_bed/1KG_Phase3.WG.CLEANED.EUR_MAF001 \
--clump quantanx_metaanalysis_filtered.gz \
--clump-p1 1 \
--clump-p2 1 \
--clump-r2 0.1 \
--clump-kb 500 \
--clump-range magma_genes_sig_plinkformat.txt \
--out quantanx_p1_rangemagma

```

```{r}

gene_loci <- fread("quantanx_p1_rangemagma.clumped.ranges.txt", data.table = FALSE)

# Genes appear in multiple clumps and need to identify where they multiple clumps share any member and merge together into one cluster. Can also look at POS to calculate this.

library(igraph)

gene_loci_sig <- gene_loci %>%
  mutate(RANGES = if_else(RANGES == "[]", NA_character_, RANGES)) %>% #PLINK represents no overlapping range as "[]" rather than NA - convert to NA
  filter(!is.na(RANGES)) %>%
  mutate(RANGES = str_remove_all(RANGES, "[\\[\\]]") %>% str_trim()) %>%
  filter(RANGES != "") %>%
  separate_rows(RANGES, sep = ",") %>%
  mutate(RANGES = str_trim(RANGES)) %>%
  filter(RANGES %in% magma_genes_sig$SYMBOL)

gene_edges <- gene_loci_sig %>%
  group_by(SNP) %>%
  filter(n() > 1) %>%          # only clumps with multiple sig genes
  summarise(genes = list(RANGES)) %>%
  mutate(edges = map(genes, ~ as.data.frame(t(combn(.x, 2))))) %>%
  unnest(edges) %>%
  rename(from = V1, to = V2)

gene_edges <- gene_edges %>% distinct(from, to)
g <- graph_from_data_frame(gene_edges, directed = FALSE)
# Add isolated nodes (sig genes that never co-occur with another sig gene)
isolated <- setdiff(magma_genes_sig$SYMBOL, V(g)$name)
g <- add_vertices(g, length(isolated), name = isolated)

components <- components(g)
sig_genes_clustered <- tibble(GENE = names(components$membership),
       cluster = components$membership) %>%
  group_by(cluster) %>%
  summarise(genes = list(sort(GENE)), n_genes = n())

sig_genes_clustered <- sig_genes_clustered %>%
  mutate(genes_str = map_chr(genes, ~ paste(.x, collapse = ", ")))

sig_genes_clustered

```

## Gene sets and gene tissue
```{r}

gene_sets <- fread("FUMA_job686083/magma.gsa.out.txt")
gene_sets <- gene_sets %>% 
  filter(P < 0.0001) %>% 
  arrange(P) %>%
  mutate(bonf_sig = if_else(P < 2.94e-06, "Yes", "No"))

gene_tissue_dev <- fread("FUMA_job686083/magma_exp_bs_dev_avg_log2RPKM.gsa.out.txt", skip = 4)
gene_tissue_dev <- gene_tissue_dev %>% 
  filter(P < 0.05/nrow(gene_tissue_dev)) %>%
  arrange(P)

gene_tissue_body <- fread("/FUMA_job686083/magma_exp_gtex_v8_ts_general_avg_log2TPM.gsa.out.txt", skip = 4)
gene_tissue_body <- gene_tissue_body %>% 
  filter(P < 0.05/nrow(gene_tissue_body)) %>%
  arrange(P)
gene_tissue_body

```

# LDSC rgs
Genetic correlations with external traits
```{bash}

output_dir="/rg_ldsc"

# My focal GWAS
focal="/munged/quantanx_metaanalysis.sumstats.gz"

# Sumstats location
dir1="/externalrg/external_sumstats/munged"
dir1_files=( #list trait names, example subset shown here
ADHD06
SCHI07
SUBD01
PPDP01
SMOK07
EXTE01
AGAB02
EXTR02
CONS02
OPEN02
IBOS01
ASTH02
BODY16
BIPO03
AUTI09
ANOR02
EDUC03
INCO01
COAD03
DIAB05
)

# Create a single comma-separated list of all external sumstats starting with focal phenotype
all_sumstats="$focal"

for f in "${dir1_files[@]}"; do
    all_sumstats="${all_sumstats},${dir1}/${f}.sumstats.gz"
done
# check with echo "$all_sumstats"

# If sumstats are in more than one location, list as above and add to all_sumstats
# for f in "${dir2_files[@]}"; do
#    all_sumstats="${all_sumstats},${dir2}/${f}.sumstats.gz"
# done

python /ldsc/ldsc.py \
    --rg "$all_sumstats" \
    --ref-ld-chr /ldsc/eur_w_ld_chr/ \
    --w-ld-chr /ldsc/eur_w_ld_chr/ \
    --out "${output_dir}/quant_anx_rg"

```

Extract results from log file
```{bash}

awk '/^p1/{found=1} found' quant_anx_rg.log | awk '{print $2, $3, $4, $6}' > rg_extracted.txt

```
