---
title: "quantanx_PRS"
output: html_document
authors: "Rujia Wang & Megan Skelton"
---

https://github.com/zhilizheng/SBayesRC

# Prepare summary statistics for SBayesRC
Created by Tian Xie 2023/06/12
```{r sbayesrc prep}

library(data.table) 
library(dplyr)
#install.packages("https://github.com/zhilizheng/SBayesRC/releases/download/v0.2.6/SBayesRC_0.2.6.tar.gz", repos=NULL, type="source")

## prepare your GWAS summary statistics to .ma file 1) using SNP ID from SBayesRC snp.info, 2) right format including 8 columns: SNP,A1,A2,freq,b,se,p,N  

## read snp.info from SBayesRC which includes information of 7 millions SNPs used for PRS
snp_info <- fread("/PRS/SBayesRC/ukbEUR_Imputed/snp.info")

## read your GWAS summary statistics 
GAD_summary <- fread("/meta_analysis/results/quantanx_sumstats.txt.gz")

snp_info <- snp_info %>%
  rename(SNP = ID,
         CHR = Chrom,
         BP = PhysPos) %>%
  select(SNP, CHR, BP)

GAD_summary <- GAD_summary %>%
  rename(SNP_GAD = SNP,
         b = BETA,
         se = SE,
         p = P,
         freq = FREQ) %>%
  mutate(A1 = toupper(A1),
         A2 = toupper(A2)) %>%
  select(SNP_GAD, CHR, BP, A1, A2, freq, b, se, p, N)

mergedData <- GAD_summary %>%
  left_join(snp_info, by=c("CHR","BP"))
mergedData <- mergedData %>%
  select(SNP, A1, A2, freq, b, se, p, N) %>%
  filter(!is.na(SNP))

fwrite(mergedData, "quantanx_cleaned_sumstats.ma", sep="\t")

```

# Generate SNP weights in SBayesRC
Set variables
```{bash set variables}

ma_file="/PRS/SBayesRC/quantanx_cleaned_sumstats.ma" # GWAS summary in COJO format (the only input - change to your summary statistics)
ld_folder="/PRS/SBayesRC/ukbEUR_Imputed"        # LD reference (downloaded from "Resources")
annot="/PRS/SBayesRC/annot_baseline2.2.txt"     # Functional annotation (downloaded from "Resources")
out_prefix="/PRS/SBayesRC/quantanx_PRS"   # Output prefix
threads=8                       # Number of CPU cores

```

Generate new SNP weights
```{bash snp weights}

export OMP_NUM_THREADS=$threads

# Tidy: optional step, tidy summary data
## "log2file=TRUE" means the messages will be redirected to a log file 
Rscript -e "SBayesRC::tidy(mafile='$ma_file', LDdir='$ld_folder', \
                  output='${out_prefix}_tidy.ma', log2file=TRUE)"
## Best practice: read the log to check issues in your GWAS summary data.  

# Impute: optional step if your summary data doesn't cover the SNP panel
Rscript -e "SBayesRC::impute(mafile='${out_prefix}_tidy.ma', LDdir='$ld_folder', \
                  output='${out_prefix}_imp.ma', log2file=TRUE)"

# SBayesRC: main function for SBayesRC
Rscript -e "SBayesRC::sbayesrc(mafile='${out_prefix}_imp.ma', LDdir='$ld_folder', \
                  outPrefix='${out_prefix}_sbrc', annot='$annot', log2file=TRUE)"

```

# Calculate PRS in PLINK
Assumes output file is SNP ID col 1, EA col 2, Effect col 3 
```{bash plink prs}

plink \
--bfile data_forSBayesRC_allChr \
--keep gad_id_list.txt \
--score ${out_prefix}_sbrc.txt 1 2 3 header sum center \
--out ${out_prefix}

```

# PRS prediction
Filter to required ancestry, ensure related individuals are removed

## Anxiety symptom severity
Using scale() such that coefficient and z-score represent a 1 SD increase in PRS.

t-value output is labelled as z-score in the table (est/SE): For large samples, the t-distribution ≈ standard normal so can be interpreted as a z-score. glm output uses z-scores, so we are doing this for consistency - most PRS predict case-control binary outcomes
```{r prs anx sympt sev}

library(data.table)
library(dplyr)
library(pscl) # for pR2()

dat <- readRDS("quantanx_prs.rds")
prs <- fread("quantanx_PRS.profile")
prs <- prs %>% 
  select(FID, IID, quant_anxprs = SCORESUM)
dat <- dat %>%
  left_join(prs, by = c("FID", "IID")) %>%
  filter(!is.na(quant_anxprs))

mod0 <- lm(gad7_avg ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = dat)

mod1 <- lm(gad7_avg ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + scale(quant_anxprs), data = dat)

summary(mod1)

# Extract R2
R2_mod0 <- summary(mod0)$r.squared
R2_mod1 <- summary(mod1)$r.squared

# Compute R2 specifically for PRS
R2_PRS <- R2_mod1-R2_mod0
print(paste("R2 for PRS:", R2_PRS))

```

## Anxiety case control
### Liability conversion functions
```{r r2 conversion function}

#Lee SH, Goddard ME, Wray NR, Visscher PM. (2012) A better coefficient of determination for genetic profile analysis. Genet Epidemiol. 2012 Apr;36(3):214-24.
  # K baseline disease risk
  # r2 from a *linear regression model* attributable to genomic profile risk score
  # P proportion of sample that are cases
  # calculates proportion of variance explained on the liability scale

h2l_R2 <- function(k, r2, p) {
  x = qnorm(1-k)
  z = dnorm(x)
  i = z/k
  C = k*(1-k)*k*(1-k)/(z^2*p*(1-p))
  theta = i*((p-k)/(1-k))*(i*((p-k)/(1-k))-x)
  h2l_R2 = C*r2 / (1 + C*theta*r2)
}

## uses the e adjustment to map Nagelkerke to the observed-scale component for ascertained case-control samples
## So, you can use the binomial model, get Nagelkerke, and then convert using the code in functions.R
## https://github.com/kn3in/ABC/blob/master/functions.R
result_h2l <- function(k, r2n, p) {
  x <- qnorm(1 - k)
  z <- dnorm(x)
  i <- z / k
  cc <- k * (1 - k) * k * (1 - k) / (z^2 * p * (1 - p))
  theta <- i * ((p - k)/(1 - k)) * (i * ((p - k) / ( 1 - k)) - x)
  e <- 1 - p^(2 * p) * (1 - p)^(2 * (1 - p))
  h2l <- cc * e * r2n / (1 + cc * e * theta * r2n)
}

```

### Regressions
```{r case control prediction}

# using pcsl functions: can also use rcompanion package with nagelkerke(model)$Pseudo.R.squared.for.model.vs.null[3,]

table(dat$anxiety_disorder, useNA = "always") # check case/control counts prior to regressions

mod0_glm <- glm(anxiety_disorder ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, family = "binomial", data = dat)

mod1_glm <- glm(anxiety_disorder ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + scale(quant_anxprs), family = "binomial", data = dat)

summary(mod1_glm)

# Extract R2
pR0 <- pR2(mod0_glm)["r2CU"]
pR1 <- pR2(mod1_glm)["r2CU"]
delta_nagelkerke <- as.numeric(pR1 - pR0)
print(paste("Observed R2 for PRS:", delta_nagelkerke))

Liab_R2 <- result_h2l(0.2, delta_nagelkerke, mean(dat$anxiety_disorder))  # assuming 20% prevalence of anxiety disorder
Liab_R2

## compare to conversion method using lm;fit linear regressions with 0/1 outcome. Should be the ~same
mod0_lm <- lm(anxiety_disorder ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 +
              PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = dat)
mod1_lm <- lm(anxiety_disorder ~ Batch + Centre + PC1 + PC2 + PC3 + PC4 +
              PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + scale(quant_anxprs),
              data = dat)
r2_obs <- summary(mod1_lm)$r.squared - summary(mod0_lm)$r.squared
h2_liability_from_lm <- h2l_R2(0.2, r2_obs, mean(dat$anxiety_disorder))
h2_liability_from_lm

```